\documentclass[twoside,a4paper,draft]{refart}

% refman:
%\documentclass[pagesize,twoside,a5paper,smallborder,10pt]{refart}
% example: https://de.sharelatex.com/project/56d7250dc2b46eb00522b9c6


% tlmgr --gui


\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage[]{hyperref}
\usepackage{comment}
\usepackage{color}
\definecolor{codebg}{gray}{0.85}
\definecolor{exmpl}{rgb}{.88,.80,.32}
\definecolor{exmplbg}{rgb}{.99,.98,.93}
\definecolor{note}{rgb}{.32,.88,.32}
\definecolor{notebg}{rgb}{.91,.98,.91}

\usepackage{listings} % https://de.sharelatex.com/learn/Code_listing
\lstMakeShortInline[basicstyle=\small\ttfamily\color{blue}]|

\lstset{keywordstyle=\sffamily\color{red},language=C}
\def\inliner{\lstinline[basicstyle=\ttfamily,language=R,keywordstyle={}]} % http://tex.stackexchange.com/questions/44702/can-you-change-lstinline-without-changing-the-global-lstset

% http://tex.stackexchange.com/questions/215520/output-from-tree-command-in-a-listing
\usepackage{dirtree} % http://tug.ctan.org/macros/generic/dirtree/dirtree.pdf
\lstdefinestyle{tree}{
    literate=
    {├}{{\smash{\raisebox{-1ex}{\rule{1pt}{\baselineskip}}}\raisebox{0.5ex}{\rule{1ex}{1pt}}}}1 
    {─}{{\raisebox{0.5ex}{\rule{1.5ex}{1pt}}}}1 
    {└}{{\smash{\raisebox{0.5ex}{\rule{1pt}{\dimexpr\baselineskip-1.5ex}}}\raisebox{0.5ex}{\rule{1ex}{1pt}}}}1 
}


\newcommand{\R}{R}
\providecommand{\specversion}{0.1}

% http://tex.stackexchange.com/questions/10085/colored-fancy-boxes
% mdframed or tcolorbox? > tcolorbox cannot compile the document anymore
% http://mirror.unl.edu/ctan/macros/latex/contrib/tcolorbox/tcolorbox.pdf
\usepackage[]{mdframed}
% ftp://ftp.rrzn.uni-hannover.de/pub/mirror/tex-archive/macros/latex/contrib/mdframed/mdframed.pdf

%\mdfdefinestyle{mystyle}{leftmargin=1cm,linecolor=blue}
%\newmdenv[frametitle=Example,style=mystyle]{example}
%\newmdenv[linecolor=note,frametitle=Note]{note}

\mdfdefinestyle{theoremstyle}{%
    linewidth=5pt,%
    frametitlerule=false,%
    topline=false,bottomline=false,rightline=false,
    innertopmargin=.2\topskip,
    frametitlebelowskip=.2\topskip,
    skipabove=\topskip,
    skipbelow=\topskip
}
\mdtheorem[style=theoremstyle,linecolor=note,backgroundcolor=notebg]{note}{Note}
\mdtheorem[style=theoremstyle,linecolor=exmpl,backgroundcolor=exmplbg]{exmpl}{Example}

\title{Bag to never differ---a standard for archiving research data and code with containers}
%\title{Bagtainer Specification}
\author{Daniel Nüst, o2r Team}
\date{February 2016\\ \vspace{1em}Version \specversion}

\begin{document}

\maketitle


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}

Bagtainers is the working title for an implementation of the idea of executable research containers (ERC) in the context of the DFG research project Opening Reproducible Research (o2r, \url{http://o2r.info}).


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Concept}

The bagtainer specification is inspired by two approaches to improve development and operation of software.

First, "convention over configuration"\footnote{\url{https://en.wikipedia.org/wiki/Convention_over_configuration}}, e.g. as practiced in the Java build tool \href{https://maven.apache.org/what-is-maven.html}{Maven}\footnote{\url{https://books.sonatype.com/mvnref-book/reference/installation-sect-conventionConfiguration.html}}.

Second, "DevOps", see Boettiger and \url{https://en.wikipedia.org/wiki/DevOps}.

In following these guidelines this specification also makes sure to fit well with the concrete standards and technologies that it is based on, namely the Internet-Draft BagIt (ref!) and Docker. We want to create a directory structure with default file names and sensible defaults for settings that require only minimal configuration in 80\% of the cases, while allowing to override each of the settings if need be and proving full customizability in the remaining 20\%.

For example, the main command to compile the text manuscript in a bagtainer could be \inliner{knitr::knit("paper.*")}, with \texttt{*} being replaced by the extension of the first file named \texttt{paper}. This covers many of the potential file formats such as \texttt{Rnw} or \texttt{Rmarkdown} with the default formats. However, if a user wants to use \inliner{rmarkdown::render(..)} on a file named \texttt{publication.md}, then the default behaviour can be overwritten.

\subsection{How to interact with a bagtainer?}

The interaction steps with a bagtainer to (re-)run the contained analysis are as follows:

\begin{enumerate}
    \item (if compressed) open the bag
    \item execute the container
    \item (automatically) compare the output contained in the bag with the just created new output
\end{enumerate}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Specification}

\subsection{Directory structure}

\dirtree{%
.1 0001.
.2 tagmanifest-md5.txt.
.2 manifest-md5.txt.
.2 bag-info.txt.
.2 bagit.txt.
.2 data.
.3 Bagtainerfile.
.3 input.
.4 Bagtainer.R.
.4 paper.Rnw.
.3 output.
.3 container.
.4 Dockerfile.
}

\subsection{Conformance\footnote{Wording as well as styling based on \url{https://w3ctag.github.io/design-principles/#conformance}}}

The key words “MUST”, “MUST NOT”, “REQUIRED”, “SHALL”, “SHALL NOT”, “SHOULD”, “SHOULD NOT”, “RECOMMENDED”, “MAY”, and “OPTIONAL” in the normative parts of this document are to be interpreted as described in \href{http://tools.ietf.org/html/rfc2119}{RFC 2119}. However, for readability, these words do not appear in all uppercase letters in this specification.

All of the text of this specification is normative except sections explicitly marked as non-normative, examples, and notes.

Examples in this specification are introduced with the words “for example” or are set apart from the normative text with class="example", like this:

\begin{exmpl*}
This is an example of an informative example.
\end{exmpl}

Informative notes begin with the word “Note” and are set apart from the normative text with class="note", like this:

\begin{note*}
Note, this is an informative note. 
\end{note}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Example use cases}




%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Tools}

\subsection{How to use a bagtainer from \R{}}

The  R package \texttt{bagtainer} provides the following commands to build, run, and compare bagtainers.

\begin{enumerate}
    \item \colorbox{codebg}{\inliner{initialize()}} creates the directory structure of the bagtainer, including the required Dockerfile with the required dependencies
    \item 
\end{enumerate}

\subsection{How to use a bagtainer from \texttt{\$}}

We could provide bash scripts to execute containers.

\subsection{How to use a bagtainer in the cloud/the o2r web applications?}


\section{Random notes}

Execute ERC from R, node, oder from bash!
Bagordner "container", per default werden alle Dateien darin als Dockercontainer gestartet und mit dem Standardverzeichnis "input" als ÜbergabeParameter gestartet

Default Backerfile enrhält die standardKommandos die überschrieben werden können, beliebige VMs können gestartet werden

Backerfile muss so heißen und ist YAML

Must a bagtainer contain license information? It should be possible to put it in the semantic metadata

Spec could contain recommendations for bagtainers, such as adding a human readable readme with info about alle the files in the data folder

What about "Geoboxes" (similar to Belmann et al's Bioboxes) ?

self-coherent bagtainers vs. tools around the containers, data outside the containers, ...
pro "data outside containers": hashes, compatible with GitBag-light, ...
 

\section{Out of scope}

writing good documentation of the included data set/analysis/workflow


\end{document}
